---
title: "Figures1.Rmd"
author: "Fayzan Chaudhry"
date: "4/28/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r }
rm(list = ls())
library(DropletUtils)
library(scater)
library(scran)
library(ggplot2)
library(Seurat)
library(dplyr)
library(viridis)
library(EnhancedVolcano)
library(formattable)
library(pheatmap)
library(ape)
library(ComplexHeatmap)
library(patchwork)
library(scales)
library(wesanderson)
library(ggrepel)
library(ggeasy)
library(data.table)
library(scuttle)
library(SingleR)
library("Seurat")
library("monocle3")
library("org.Hs.eg.db")
library("dplyr")
addTaskCallback(function(...) {set.seed(100);TRUE})
```
#Load object
```{r }
combined_sce1 <- readRDS("FinalObj.RDS")
combined_sce2 <- combined_sce1
sce2 <- as.Seurat(combined_sce2)
Idents(sce2) <- sce2$clusters
#
sce1 <- sce2
sce1
```

##############
#UMAPs For Figure 5A and 5B
```{r }
color_samp <- wes_palette("Darjeeling1",3)
wt_col <- color_samp[2]
ko_col <- color_samp[3]
my_color_palette <- hue_pal()(length(levels(sce1$clusters)))
DimPlot(sce1) + ggtitle("UMAP by Clusters")
DimPlot(sce1, group.by = "condition", cols = c(wt_col, ko_col), label = FALSE) + ggtitle("UMAP by WT vs KO") + theme(legend.text=element_text(size=13))
```
#MAKE HEAT MAP, CLUSTERPLOTS, VLN
```{r }
markers <- FindAllMarkers(sce1, min.pct = 0.15, min.diff.pct = 0.04, logfc.threshold = .25, test.use = "MAST") #Try with #MAST
top20 <- markers %>%  group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC)
sce1 <- ScaleData(object = sce1, block.size=100)
temp3 <- GetAssayData(sce1, slot = "scale.data")
mat<- temp3[unique(c("JUN", top20$gene)), ] %>% as.matrix()
heatmapanno <- HeatmapAnnotation(Group = sce1$condition, Cluster = Idents(sce1), col = list(Group=c("WT" = wt_col, "KO" = ko_col), Cluster =c("1" = "#F8766D", "2" =  "#C49A00", "3" =  "#53B400", "4" =  "#00C094", "5" =  "#00B6EB", "6" =  "#A58AFF", "7" ="#FB61D7")))
col_fun = circlize::colorRamp2(c(-1, 0, 2), c("#140b34" , "#84206b", "#f6d746"))
```
##Figure 5D 
```{r }
Heatmap(mat, name = "Scaled Expression",  
        column_split = Idents(sce1),
        #column_title = "Cluster",
        cluster_columns = FALSE, 
        show_column_dend = FALSE,
        cluster_column_slices = FALSE,
        column_title_gp = gpar(fontsize = 7),
        column_gap = unit(.6, "mm"),
        cluster_rows = TRUE,
        show_row_dend = TRUE,
        row_dend_width = unit(20, "mm"),
        col = col_fun,
        row_names_gp = gpar(fontsize = 8),
        column_title_rot = 0,
        top_annotation = heatmapanno,
        show_column_names = FALSE,
        use_raster = TRUE,
        raster_quality = 4)
```
#Figure 5F
sce3 <- subset(sce1, ident = 7, invert = TRUE)
v1 <- VlnPlot(sce3, "BAX", split.by = "condition", split.plot = TRUE, cols = c(wt_col, ko_col), pt.size = 0)
v2 <- VlnPlot(sce3, "TNFRSF12A", split.by = "condition", split.plot = TRUE, cols = c(wt_col, ko_col),  pt.size = 0)
v3 <- VlnPlot(sce3, "JUN", split.by = "condition", split.plot = TRUE, cols = c(wt_col, ko_col),  pt.size = 0)
v1+v2 + v3

##Figure 5E
```{r }
Idents(sce1) <-sce1$condition 
markers1 <- FindAllMarkers(sce1, min.pct = 0.05, min.diff.pct = 0.02, logfc.threshold = .05, test.use = "MAST")
top <- markers1 %>% dplyr::filter(cluster=="WT")
keyvals <- ifelse(
  top$avg_log2FC < -.10, '#F2AD00',
  ifelse(top$avg_log2FC > .10, '#00A08A',
         'grey50'))
keyvals[is.na(keyvals)] <- 'grey50'
names(keyvals)[keyvals == '#00A08A'] <- 'Wildtype'
names(keyvals)[keyvals == 'grey50'] <- 'NS'
names(keyvals)[keyvals == '#F2AD00'] <- 'Knockout'
apotosis1 <-c("FAS",
              
              "TNFRSF10A",
              "TNFRSF10B",
              "TNFRSF10C",
              "TNFRSF10D",
              "TNFRSF11B",
              "TNFSF10",
              "TNFRSF1A",
              "TNFRSF12A",
              "FADD",
              "CFLAR",
              "CASP1",
              "CASP2",
              "CASP3",
              "CASP4",
              
              "CASP6",
              "CASP7",
              "CASP8",
              "CASP9",
              "CASP10",
              "CASP14",
              "NAIP",
              "BIRC2",
              "BIRC3",
              "XIAP",
              "BIRC5",
              "BIRC6",
              "BIRC7",
              "BCL2",
              "MCL1",
              "BCL2L1",
              "BCL2L2",
              "BCL2A1",
              "BCL2L10",
              "BAX",
              "BAK1",
              "BOK",
              "BID",
              "BCL2L11",
              "BMF",
              "BAD",
              "BIK",
              "HRK",
              "PMAIP1",
              "BNIP3",
              "BNIP3L",
              "BCL2L14",
              "BBC3",
              "BCL2L12",
              "BCL2L13",
              "APAF1",
              "CYCS",
              "DIABLO",
              "HTRA2",
              "AIFM1",
              "ENDOG",
              "CARD8",
              "CARD6",
              "NOX5",
              "TP53",
              "CDKN1A",
              "CDKN1B",
              "CDKN2A",
              "CDKN2B")
ofinterest <- c("MKI67",
                "TOP2A",
                "NES",
                "MAP2",
                "TH",
                "EN1",
                "NR4A2",
                "HIF1A",
                "TP53",
                "RELA",
                "NFKB1",
                "NFKB2",
                "JUN",
                "FOS", "LMO3", "HES5")
geneofinterest <- c(ofinterest, apotosis1)
samplabels1 <- c("PEG10", "DYNC1I2", "JUN",
                 
                 "NMT1",
                 "PPDPF",
                 "PHPT1",
                 "NDUFA13",
                 "NDUFC2",
                 "NDUFA11",
                 "NME2",
                 "NDUFA12",
                 "RPS27L",
                 "INSM1",
                 "FABP5",
                 "HSPA1A",
                 "WLS",
                 "MICOS13",
                 "ASCL1",
                 "SOX2")
sce4 <- subset(sce1, ident == 3 | ident == 5 | ident == 6)
Idents(sce4) <-sce4$condition
markers2 <- FindAllMarkers(sce4, min.pct = 0.01, min.diff.pct = 0.01, logfc.threshold = .05, test.use = "MAST")
top2 <- markers2 %>% dplyr::filter(cluster=="WT")
keyvals <- ifelse(
  top2$avg_log2FC > .10 & top2$p_val_adj < .01, '#00A08A',
  ifelse(top2$avg_log2FC < -.10 & top2$p_val_adj < .01, '#F2AD00',
         'grey50'))
keyvals[is.na(keyvals)] <- 'grey50'
names(keyvals)[keyvals == '#00A08A'] <- 'Wildtype'
names(keyvals)[keyvals == 'grey50'] <- 'NS'
names(keyvals)[keyvals == '#F2AD00'] <- 'Knockout'
samplabels<- c("RPS27L", "PHLDA3", "TUBB2A","TUBB2B","RPL27A","PHPT1","CDKN1A","BAX",
      "NMT1",
      "BBC3",
      "CDKN2B",
      "TNFRSF10B")
EnhancedVolcano(top2,
                lab = top2$gene,
                x = 'avg_log2FC',
                y = 'p_val_adj',
                title = 'WT vs KO Top DE Genes',
                pCutoff = .01,
                FCcutoff = .10,
                pointSize = 1.0,
                labSize = 3.3,
                selectLab = samplabels,
                col=c('blue', 'green', 'black', 'red3', "yellow", "brown", "grey"),
                colAlpha = .8,
                colCustom = keyvals,
                drawConnectors = TRUE,
                widthConnectors = 0.1,
                lengthConnectors = unit(0.01, "npc"),
                cutoffLineType = "dotted", 
                cutoffLineCol = "#00000066",
                #titleLabSize = 10,
                #subtitleLabSize = 8,
                #captionLabSize = 8
)  + xlim(-.75, .75) 
```
##Barplot and Gene plot 5SE 5SD 5SB
```{r }
table2 <- as.matrix(table(sce1@meta.data$condition, sce1@meta.data$cluster))
table2[1,] <- table2[1,] * table(sce1$condition)[2]/table(sce1$condition)[1]
table3 <- data.table(t(t(table2) / colSums(table2)))
colnames(table3) <- c("Condition", "Cluster", "Fraction")
ggplot(data=table3, aes(x=Cluster, y=Fraction, fill=Condition)) +
  geom_bar(stat="identity") + scale_fill_manual(values = c(ko_col, wt_col)) + ggtitle("Clusters by Condition")
sce1 <- sce2 
Idents(sce1) <- sce1$condition
sce1 <- subset(x = sce1, idents = "KO") 
th <- length(WhichCells(sce1, slot = 'counts', expression = TH > 0)) /  length(WhichCells(sce1, slot = 'counts'))
mki67 <- length(WhichCells(sce1, slot = 'counts', expression = MKI67 > 0)) /  length(WhichCells(sce1, slot = 'counts'))
map2 <- length(WhichCells(sce1, slot = 'counts', expression = MAP2 > 0)) /  length(WhichCells(sce1, slot = 'counts'))
all_ko <- c(map2, th, mki67)
sce1 <- sce2
Idents(sce1) <- sce1$condition
sce1 <- subset(x = sce1, idents = "WT")
th <- length(WhichCells(sce1, slot = 'counts', expression = TH > 0)) /  length(WhichCells(sce1, slot = 'counts'))
mki67 <- length(WhichCells(sce1, slot = 'counts', expression = MKI67 > 0)) /  length(WhichCells(sce1, slot = 'counts'))
map2 <- length(WhichCells(sce1, slot = 'counts', expression = MAP2 > 0)) /  length(WhichCells(sce1, slot = 'counts'))
all_wt <- c(map2, th, mki67)
temp <- data.table(
  Gene = c("MAP2", "TH", "MKI67"),
  Group =c("WT", "WT", "WT", "KO", "KO", "KO"),
  Values = c(all_wt, all_ko)
)

ggplot(temp, aes(x = Gene, y = Values, fill = Group)) + 
  geom_bar(stat = "identity", position = "dodge") + ggtitle("Fraction of Cells Expressing Each Gene") + ggeasy::easy_center_title() +
  theme(plot.title = element_text(face="bold")) +theme(text=element_text(size=8)) + scale_fill_manual(values = c(ko_col, wt_col)) 
```
##Supplement 5C
```{r }
sce1 <- sce2
sce1.list = list(sce1[, sce1@meta.data$condition == "WT"], sce1[, sce1@meta.data$condition == "KO"])
f1 <-FeaturePlot(sce1.list[[1]], feature="MAP2", order=T, pt.size=1.0) + ggplot2::labs(title="MAP2 (WT)") + scale_colour_gradientn(colours = terrain.colors(10))
f2 <-FeaturePlot(sce1.list[[2]], feature="MAP2", order=T, pt.size=1.0) + labs(title="MAP2 (KO)") + scale_colour_gradientn(colours = terrain.colors(10))
#PBX1
f3 <-FeaturePlot(sce1.list[[1]], feature="PBX1", order=T, pt.size=1.0) + labs(title="PBX1 (WT)") + scale_colour_gradientn(colours =terrain.colors(10))
f4 <-FeaturePlot(sce1.list[[2]], feature="PBX1", order=T, pt.size=1.0) + labs(title="PBX1 (KO)") + scale_colour_gradientn(colours = terrain.colors(10))
#MKI67
f5 <-FeaturePlot(sce1.list[[1]], feature="MKI67", order=T, pt.size=1.0) + labs(title="MKI67 (WT)") + scale_colour_gradientn(colours = terrain.colors(10))
f6 <-FeaturePlot(sce1.list[[2]], feature="MKI67", order=T, pt.size=1.0) + labs(title="MKI67 (KO)") + scale_colour_gradientn(colours = terrain.colors(10))
#SERT
f7 <-FeaturePlot(sce1.list[[1]], feature="SLC6A4", order=T, pt.size=1.0) + labs(title="SLC6A4 (WT)") + scale_colour_gradientn(colours = terrain.colors(10))
f8 <-FeaturePlot(sce1.list[[2]], feature="SLC6A4", order=T, pt.size=1.0) + labs(title="SLC6A4 (KO)") + scale_colour_gradientn(colours = terrain.colors(10))
#TPH1
f9 <-FeaturePlot(sce1.list[[1]], feature="TPH1", order=T, pt.size=1.0) + labs(title="TPH1 (WT)") + scale_colour_gradientn(colours = terrain.colors(10))
f10 <-FeaturePlot(sce1.list[[2]], feature="TPH1", order=T, pt.size=1.0) + labs(title="TPH1 (KO)") + scale_colour_gradientn(colours = terrain.colors(10))

(f1/f2) | (f3/f4) | (f5/f6) 
```

##FIGURE 5C
```{r }
#load cellline data
cellline <- readRDS("cellline_clean.rds")
cellline <- as.Seurat(cellline)

#La Manno dataset used a reference
combined_sce1 <- as.SingleCellExperiment(sce1)
combined_sce1$clusters <- sce1$clusters
combined_sce_lamanno <- scRNAseq::LaMannoBrainData("human-embryo")
# SingleR() expects reference datasets to be normalized and log-transformed.
combined_sce_lamanno <- logNormCounts(combined_sce_lamanno) 
predictions2 <- SingleR(test=combined_sce1, assay.type.test=2, 
                        ref=combined_sce_lamanno, labels =combined_sce_lamanno$Cell_type, de.method="wilcox", clusters=combined_sce1$clusters)
predictions2$pruned.labels #Didn't assign
sce5 <- as.Seurat(combined_sce1)
Idents(sce5) <- sce5$clusters
sce5 <- RenameIdents(object = sce5, `1` = "hNbM", `2` = "hNbM", `3` = "hProgFPL", '4'= "hNbM", "5" = "hProgFPL", "6" = "hProgFPL", "7" = "hPeric")
DimPlot(sce5) + ggtitle("UMAP by Annotation")
combined_sce_lamanno1  <- subset(combined_sce_lamanno , ,Cell_type == c('hNbM') | Cell_type == c('hNbML1') | Cell_type == c('hProgFPL'))
predictions3 <- SingleR(test=combined_sce1, assay.type.test=2, 
                        ref=combined_sce_lamanno1, labels =combined_sce_lamanno1$Cell_type, de.method="wilcox")
combined_sce1$pruned.labels <- predictions3$pruned.labels
sce6 <- as.Seurat(combined_sce1)
v1 <- VlnPlot(sce6, "BAX", group.by = "pruned.labels", cols = c(wt_col, ko_col), pt.size = 0, split.by = "condition")
v2 <- VlnPlot(sce6, "TNFRSF12A", group.by = "pruned.labels",  cols = c(wt_col, ko_col),  pt.size = 0, split.by = "condition")
v3 <- VlnPlot(sce6, "JUN", group.by = "pruned.labels", cols = c(wt_col, ko_col),  pt.size = 0, split.by = "condition")
v1 + v2 + v3

cellline <- readRDS("cellline_clean.rds")
cellline <- as.Seurat(cellline)

f1 <- FeaturePlot(sce1, feature="HES5", order=T, pt.size=1.0) + labs(title="HES5") + scale_colour_gradientn(colours = terrain.colors(10))
f1 <- FeaturePlot(sce1, feature="HES5", order=T, pt.size=1.0) + labs(title="HES5") + scale_colour_gradientn(colours = terrain.colors(10))

#ALTERNATE FIGURE REQUESTS
sce4 <- subset(sce1, ident == 1 |ident == 2 | ident == 4)
Idents(sce4) <-sce4$condition

markers2 <- FindAllMarkers(sce4, min.pct = 0.01, min.diff.pct = 0.01, logfc.threshold = .05, test.use = "MAST")
top2 <- markers2 %>% dplyr::filter(cluster=="WT")
keyvals <- ifelse(
  top2$avg_log2FC >= .10 & top2$p_val_adj < .01, '#00A08A',
  ifelse(top2$avg_log2FC <= -.10 & top2$p_val_adj < .01, '#F2AD00',
         'grey50'))
keyvals[is.na(keyvals)] <- 'grey50'
names(keyvals)[keyvals == '#00A08A'] <- 'Wildtype'
names(keyvals)[keyvals == 'grey50'] <- 'NS'
names(keyvals)[keyvals == '#F2AD00'] <- 'Knockout'
samplabels<- c("PEG10",
               "NMT1",
               "HSPA8", "RPS27L",
               "NDUFA12",
               "DYNLT1",
               "PDCD5",
               "PHPT1",
               "KRTCAP2", "SYT14",
               "SOX2",
               "HSPA1B",
               "HSPA1A","HSPA2A","DNAJB1","HSPA8")
EnhancedVolcano(top2,
                lab = top2$gene,
                x = 'avg_log2FC',
                y = 'p_val_adj',
                title = 'WT vs KO Top DE Genes',
                pCutoff = .01,
                FCcutoff = .10,
                pointSize = 1.0,
                labSize = 3.3,
                selectLab = samplabels,
                col=c('blue', 'green', 'black', 'red3', "yellow", "brown", "grey"),
                colAlpha = .8,
                colCustom = keyvals,
                drawConnectors = TRUE,
                widthConnectors = 0.1,
                lengthConnectors = unit(0.01, "npc"),
                cutoffLineType = "dotted", 
                cutoffLineCol = "#00000066",
                #titleLabSize = 10,
                #subtitleLabSize = 8,
                #captionLabSize = 8
)  + xlim(-.8, .8) 
```



---
title: Figures1.R
author: fayzan
date: '2024-04-29'

---
